Untwisting the Tetrahelix (v0.5)

### 2017-02-22 doc/TetrahelixUntwist.tex
chirality understood better, now must code

From Robert Gray's site, repeating formula by H.S.M. Coxeter:
\begin{equation}
  \label{graycoxeter}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta) $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession on every colored rail.

@@

We seek a formula to generate optimal tetrahelices that accepts a
parameter that allows us to choose the tetrahelix conveniently. The
pitch of the helix is an obious choice, but is not defined when the
curvature is $0$, and important special case. The radius or the axial
distance between two nodes on the same rail are obvious choices, but
perhaps the clearest choice is to build formula that takes as its
input the ``rail angle'' $\rho$. We define $\rho$ to be the angle
formed in the X,Y plane $\angle R_i O R_{i+1}$ projecting out the $z$
axis and sighting along the positive $z$ axis. In other words, $\rho$
controls how far a rail edge of a tetrahelix deviates from being
parallel with the axis, or the ``twistiness'' of tetrahelix. Ideally
we will treat a positive angle as creating a clockwise tetrahelix and
a negative as creating a counter-clockwise helices.

@@

Rewriting our formulation in terms of $\rho$:
\begin{equation}
H_{general}(n,c,\rho,d_{\rho},r_{\rho}) =
\left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\rho \kappa + s\cdot c 2 \pi /3) $\\
   $ r_{\rho}  \sin(\rho \kappa + s\cdot c 2 \pi /3) $\\
   $ d_{\rho} \kappa $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
  $\kappa = n + c/3$ \\
    $s = \text{sign}{(\rho)}$ \\  
  \end{tabular}      
\end{equation}

@@

\begin{align*}
  d_{opt}^2 &= 1 - 4 \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }}^2 (\sin{ \rho / 2})^2   \\
  d_{opt}^2 &= 1 - 4 \frac{4(\sin{ \rho / 2})^2}{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16(\sin{ \rho / 2})^2}{9 ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16 sin^2(\rho / 2) }{ cos(\rho) + 9( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + 8 }   \\
    d_{opt} &= \sqrt{1 - \frac{16 sin^2(\rho / 2) }{ cos(\rho) + 9( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + 8 }}    \numberthis  \label{dopt}  \\      
\end{align*}
Therefore, by computing $r_{opt}$ as a function of $\rho$ from this equation, we can construct minimax optimal tetrahelix for an $-\rho_{bc} \leq \rho \leq \rho_{bc}$.

@@

@@

 Negative values of $\rho$ generate clockwise tetrahelices; positive generate counter-clockwise helices.
 (Note: In fact there is something wrong with my math, negative rho values do produces opposit chirality
 but the lengths are not the same as the positive values.  I need to track this down.)

@@

>>>>

From Robert Gray's site, repeating formula by H.S.M. Coxeter, for a counter-clockwise BC Helix:
\begin{equation}
  \label{graycoxeter}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta) $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession on every colored rail.

@@

We seek a formula to generate optimal tetrahelices that accepts a
parameter that allows us to choose the tetrahelix conveniently. The
pitch of the helix is an obious choice, but is not defined when the
curvature is $0$, and important special case. The radius or the axial
distance between two nodes on the same rail are obvious choices, but
perhaps the clearest choice is to build formula that takes as its
input the ``rail angle'' $\rho$. We define $\rho$ to be the angle
formed in the X,Y plane $\angle R_i O R_{i+1}$ projecting out the $z$
axis and sighting along the positive $z$ axis. In other words, $\rho$
controls how far a rail edge of a tetrahelix deviates from being
parallel with the axis, or the ``twistiness'' of tetrahelix. We use
the parameter $\chi = 1$ to indicate a chirality or counter-clockwise,
and $\chi = -1$ for clockwise.

@@

Rewriting our formulation in terms of $\rho$:
\begin{equation}
H_{general}(\chi,n,c,\rho,d_{\rho},r_{\rho}) =
\left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\chi \cdot (\rho \kappa + s\cdot c 2 \pi /3)) $\\
   $ r_{\rho}  \sin(\chi \cdot (\rho \kappa + s\cdot c 2 \pi /3)) $\\
   $ d_{\rho} \kappa $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
  $\kappa = n + c/3$ \\
    $\chi \in \{-1,1\}$ \\  
  \end{tabular}      
\end{equation}

@@

\begin{align*}
  d_{opt}^2 &= 1 - 4 \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }}^2 (\sin{ \rho / 2})^2   \\
  d_{opt}^2 &= 1 - 4 \frac{4(\sin{ \rho / 2})^2}{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16(\sin{ \rho / 2})^2}{9 ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16 sin^2(\rho / 2) }{ cos(\rho) + 9( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + 8 }   \\
    d_{opt} &= \sqrt{1 - \frac{16 sin^2(\rho / 2) }{ cos(\rho) + 9( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + 8 }}    \numberthis  \label{dopt}  \\      
\end{align*}
Therefore, by computing $r_{opt}$ as a function of $\rho$ from this equation, we can construct minimax optimal tetrahelix for an $0 \leq \rho \leq \rho_{bc}$.

@@

Note that the equitetrabeam has chirality, which becomes important in our attempt to build a
continuum of tetrahelices.

@@

 Because the equitetrabeam which as a rail angle of $0$ still has chirality, that is,
 one still must decide to connect with one-hop to the clockwise or counter-clockwise node,
 it is not poosible to build a smooth continuum where $\rho$ transitions from positive to
 negative which remains optimal. One can use a negative $\rho$ in $H_{general}$ but it does
 not produce minimax optimal tetrahelices. In other words, untwisting a conter-clockwise
 tetrahelix to $0$ and then going further does produce a clockwise tetrahelix, but one in which
 the one-hop and two-hop lengths in the wrong places (that is, two-hop becomes shorter than one-hop.)

@@

http://bit-player.org/2013/tetrahedra-with-a-twist
